---
output:
  word_document: default
  html_document: default
  pdf_document: default
---

This is data from yahoo finance.
```{r}
library(readr)
df=read.csv("AMZN.csv")
head(df,10)

```
```{r}
df1=subset(df[2])
m1 <- ts(data = df1,start = 2010, end = 2017,frequency=100)
plot.ts(m1,ylab = 'Amazon stock price',xlab='Time')
```
This analysis can obtain trends of amazon stock price. We can use time series to analyze the trends of stock price level and predict the future stock price.
```{r}
df2=read.csv("monthly-beer-production-in-austr.csv")
head(df2,10)
```
```{r}

df22=subset(df2[2],start = 1981, end = 1982)
m2 <- ts(data = df22)
plot.ts(m2,ylab = 'Monthly.beer.production',xlab='Time')
```
source: https://www.kaggle.com/datasets/shenba/time-series-datasets. This analysis can obtain trends of monthly beer production in Austrilia. We can use time series to analyze the trends of production and predict the future production trends.


Q2
1)
```{r}

TT <- 50
set.seed(1)
p <- .25
zt1 <- 1+rbinom(TT+1,1,p)
zt2 <- 1+rbinom(TT+1,1,p)
xt1 <- zt1[2:(TT+1)] * zt1[1:TT]
xt2 <- zt2[2:(TT+1)] * zt2[1:TT]

par(mfrow = c(1, 2))
plot.ts(xt1,ylab = 'x_t',xlab='t')
plot.ts(xt2,ylab = 'x_t',xlab='t')

```
2)
$E(X_t)=E(Z_t*Z{t-1})=E(Z_t)*E(Z_{t-1})=E(Z_t)^2= (1*3/4+2*1/4)^2=25/16=1.5625$
$E(X_t^2)=E[Z_t^2* Z_{t-1}^2]=E[Zt^2]*E[Z_{t-1}^2]=E[Z_t^2]=49/16=3.0625$
```{r}
mean(xt1)
mean(xt2)
mean(xt1^2)
mean(xt2^2)



```
Close to the theoretical result. The xt^2 is more different. 

3)
$E(X_t*X_{t−1})=E(X_t)*E(X_{t-1})=E(Z_t*Z_{t−1})*E(Z_{t-1}*Z_{t−2})= 2.734375$
$Corr(X_t, X_{t−1})=(E(X_t*X_{t−1}) − E(X_t)E(X_{t−1})/((E(X_t^2) − (E(X_t)^2)*(E(X_{t−1}^2 − (E_{Xt−1})^2)$
$=(625/16 - (25/16)^2) / \sqrt((49/16 - (25/16)^2)(49/16 - (25/16)^2))=0.4717$
```{r}

mean(xt1[2:50]*xt1[1:49])
mean(xt2[2:50]*xt2[1:49])


cor(xt1[2:50],xt1[1:49])
cor(xt2[2:50],xt2[1:49])

```

The value are all close to theoretical result. Xt1 has closer $E(X_t*X_{t−1})$, $x_t2$ closer correlation 
4)$X_t$ can be 1,2,4
$P(x_t=1)=3/4*3/4=9/16=0.5625$
$P(x_t=2)=3/4*1/4*2=6/16=0.375$
$p(x_t=4)=1/4*1/4=1/16=0.0625$
Then we can get
$E(X_t) = 1 × 9/16 + 2 × 6/16 + 4 ×1/16= 17/16=1.0625$
$E(X_t^2) = 1 × 9/16 + 2*2 × 6/16 + 4*4 ×1/16= 41/16=2.5625$
The number we get in realizations are all higher than empirical value. Xt1 is closer to the value but still not so close to empirical value. 
```{r}
c(sum(xt1==1)/TT,sum(xt1==2)/TT,sum(xt1 == 4)/TT)
c(sum(xt2==1)/TT,sum(xt2==2)/TT,sum(xt2 == 4)/TT)
```
The probability for realization 2 is closer  

Q3Suppose a time series model

$X_t=m_t+10Z_t$ with IID noise $Z_t$ with $Z_t~N (0,1) and T=50, m_t=t^2$
.
```{r}
set.seed(12345)
tt <- seq(1,TT) 
mt <- tt 
zt3 <- rnorm(TT,0,1) 
zt4 <- rnorm(TT,0,1) 
xt3 <- mt + 10*zt3
xt4 <- mt + 10*zt4
par(mfrow = c(1, 2))

plot.ts(xt3,ylab = 'x_t',xlab='t')
plot.ts(xt4,ylab = 'x_t',xlab='t')
```
$E(X_t)=E(10Z_t+m_t)=E(m_t)+E(10Z_t)=t^2$   t=1,2,...T
$E(Xt^2)=E(100Z_t^2)+E(m_t^2*t^2)+20E(m_t*Z_t)=t^4$ t=1,2,...T
```{r}
#empirical value
mean(xt3)
mean(xt4)
mean(xt3^2)
mean(xt4^2)
```

```{r}
#empirical value
 mean(xt3[-1]*xt3[-TT])
  cor(xt3[-1],xt3[-TT])

mean(xt4[-1]*xt4[-TT])
  cor(xt4[-1],xt4[-TT])
```
E(Xt*Xt−1)=E(mt+10Zt)*E((mt-1)+(10Zt-1))=E$m_tm_{t-1}+10m_tZ_{t-1}+10Z_t*m_{t-1}+100Z_tZ_{t-1}=t^2(t-1)^2$
$Corr(X_t, X_{t−1})=0$
we cannot average over time to represent the expected value as values are dependent on t

