---
title: "Stor 556 hw 7"
output:
  word_document: default
  html_document:
    df_print: paged
---

#Q1
#a)
```{r}
library(forecast)
# Generate x_t
set.seed(1)
phi1 = 2*(1/1.2)*cos(pi/8)
phi2 = -1/1.22
x = arima.sim(list(order = c(2,0,0), ar = c(phi1,phi2) ), n = 200)
# Regression on cos() and sin()
t = time(x)
hr_1 = lm(x~1+cos(2*pi*20/200*t)+sin(2*pi*20/200*t))
# Verify the equation for k=20
a_0 = mean(x)
a_20 = 2/200*sum(x*cos(2*pi*20/200*t))
b_20 = 2/200*sum(x*sin(2*pi*20/200*t))
rbind(equation=c(a_0, a_20, b_20), coef = summary(hr_1)$coef[,1])
```
The values of a0, a20,b20 are same when k = 20.

#b
```{r}
a = rep(0, 99)
b = rep(0, 99)
for(i in 1:99){
a[i] = 2/200*sum(x*cos(2*pi*i/200*t))
b[i] = 2/200*sum(x*sin(2*pi*i/200*t))
}
a_100 = mean(x*cos(pi*t))
# Verify equations
rbind(Equation_3 = c(LHS = mean(x^2), RHS = a_0^2+0.5*sum(a^2+b^2)+a_100^2),Equation_4 = c(LHS = mean((x-mean(x))^2), RHS = 0.5*sum(a^2+b^2)+a_100^2))
```
The answers are the same

#c
```{r}
# The raw spectrum
spec.x = spec.pgram(x,log='no',taper=0,pad=0,fast=FALSE,demean=TRUE,detrend=FALSE)
I_wk = spec.x$spec
mean((x-mean(x))^2)
2/200*sum(I_wk[1:99])+1/200*I_wk[100]
```
answers are the same
#d
```{r}
k_max = which.max(I_wk)
k_max
spec.pgram(x,log='no',taper=0,pad=0,fast=FALSE,demean=TRUE,detrend=FALSE)
abline(v=spec.x$freq[15], lty = 2, lwd = 2, col = "red")
legend(0.1, 60, legend = c("w_15 = 0.075"), col = c("red"), box.lty=0)

```
#e

```{r}
plot.ts(x)
curve(a_0+a[15]*cos(2*pi*15/200*x)+b[15]*sin(2*pi*20/200*x), add = T, col = "red")

0.5*(a[15]^2 + b[15]^2)/mean((x-mean(x))^2)
```
22.39564% of the variability is explained according to eq.(4)

#Q2
#a
```{r}
library(haven)
wave <- read.delim("wave.txt", header=FALSE)
plot(ts(wave, frequency=10))
spec.pgram(wave,log='no',taper=0,pad=0,fast=FALSE,demean=TRUE,detrend=FALSE)

```
```{r}
spec.x = spec.pgram(wave,log='no',taper=0,pad=0,fast=FALSE,demean=TRUE,detrend=FALSE, spans=12)
```

#b
```{r}
freq = spec.x$freq[which.max(spec.x$spec)]
freq

```
```{r}
period = 0.1/freq
period
```

The period is 0.72s

#c
$\ln \left(\mathrm{I}\left(w_k\right)\right)=\ln \left(a w^{-5} e^{-b w^{-4}}\right)=\ln (a)-5 \ln (w)-b w^{-4}$
```{r}
a = 3.5e5
b = 0.43
k = 1:(396/2)
w = 2*pi*k/396
PMspectrum = a*w^(-5)*exp(-b*w^(-4))
spec.pgram(wave,log='no',taper=0,pad=0,fast=FALSE,demean=TRUE,
detrend=FALSE, spans=12, ylim=c(0, 400000))
lines(w/2/pi,PMspectrum, col="red")
legend(0.35, 400000, legend=c("Spectrum from a", "Pierson-Moskowitz"),
col=c("black", "red"),lty=c(1,1), cex=0.8)

```
a = 350000, b = 0.43

#Q3
```{r}
library(signal)
n = 5
low = 0.1
high = 0.5
bf = butter(n,c(low,high))
w = seq(0,pi,0.01)
numerator = 0
denominator = 0
for (j in 0:(2*n)){
j
numerator = numerator + bf$b[j+1]*exp(-j*w*1i)
denominator = denominator + bf$a[j+1]*exp(-j*w*1i)
}
h_w = Mod(numerator)^2/Mod(denominator)^2
plot(w, h_w, type="l")
abline(v=low*pi, col="red")
abline(v=high*pi, col="red")
```
The plot above demonstrates that h(w) approaches 1 for frequencies falling within the range of 0.1π to 0.5π. Such a filter is known as a band-pass filter because it permits frequencies within a specific range to pass while blocking those outside of the range.
#Q4
#a
```{r}
set.seed(1)
p4sim = arima.sim(list(order = c(1,0,0), ar = 0.5), n = 200)
plot(p4sim)

```
#b
```{r}
I = spec.pgram(p4sim,log='no',taper=0,pad=0,fast=FALSE,demean=TRUE,detrend=FALSE)
phi = seq(-1, 1, 0.01)
losses = 0
for (j in 1:100){
w_j = 2*pi*j/200
f_w = Mod(1-phi*exp(-1i*w_j))^(-2)/(2*pi)
loss = log(f_w)+I$spec[j]/(2*pi*f_w)
losses = losses + loss
}
plot(phi,losses, type='l')
```
#c
```{r}
f = function (phi){
losses = 0
for (j in 1:100){
w_j = 2*pi*j/200
f_w = Mod(1-phi*exp(-1i*w_j))^(-2)/(2*pi)
loss = log(f_w)+I$spec[j]/(2*pi*f_w)
losses = losses + loss
}
losses
}
phi_min <- optimize(f, c(-1, 1), tol = 1e-5)
phi_min
```

ϕ = 0.4894 is close to 0.5



