---
title: "HW9"
output:
  word_document: default
  html_document:
    df_print: paged
---

#Q1

#a)

$$\Delta x_{1, t}=x_{1, t}-x_{1, t-1}=\phi x_{1, \pi-1}+c x_{2, t-1}-c \phi x_{2, t, 2}+z_{1, c}-c x_{2, t-2}-Y_{1, t-1}$$


$$
=(\phi-1) x_{1, t-1}+c x_{2, t-1}-c \phi x_{2, t-2}+z_{1, t}
$$
$$
=(\phi-1) x_{1, t-1}+(c-c \phi) x_{2, t-1}+\left(c \phi x_{2, t 1}-c \phi x_{2, t 2}\right)+z_{1, t}
$$

$$
=(\phi-1) x_{1, t-1}+(c-c \phi) x_{2, t-1}+c \phi \Delta x_{2, t-1}+z_{1, t}
$$
and $\beta_1=\phi-1$, $\beta_\nu=c-c \phi$, $\gamma_1=c \phi$

#b
$$x_{1, t}=\phi X_{1, t-1}+c X_{2, t-1}-c \phi x_{2, t-2}+z_{1, t}$$
$$
x_{2, t}=x_{2, t-1}+z_{2 t}
$$

$$
x_t=\Phi_1 x_{t-1}+\Phi_2 x_{t-2}+z_t
$$
$$
\Phi_1=\left(\begin{array}{cc}
\phi & c \\
0 & 1
\end{array}\right) \quad \quad \Phi_2=\left(\begin{array}{cc}
0 & -c \phi \\
0 & 0
\end{array}\right)
$$

$$
\text { VECM: } X_t-x_{t-1}=\left(\Phi_1+\Phi_2-I\right) x_{t-1}-\Phi_2\left(x_{t-1}-X_{t-2})+z t \right.
$$
$$
\pi=\left(\begin{array}{cc}
\phi-1 & c-c \phi \\
0 & 0
\end{array}\right)
$$

$$
\widetilde{\Gamma}=\left(\begin{array}{ll}
0 & c \phi \\
0 & 0
\end{array}\right)
$$
$$
a=\left[\begin{array}{c}
\phi-1 \\
0
\end{array}\right]
$$

$$
b=\left[\begin{array}{ll}
1 & -c
\end{array}\right]
$$


#c

$$
X_{t-}-X_{t-1}=\left(\Phi_1-I\right) X_{t-1}+\Phi_2 X_{t-2}+Z_t 
$$

$$
=\left(\Phi_1-I\right) x_{t-1}-\left(\Phi_1-I\right) x_{t-2}+\left(\Phi_1-I\right) x_{t-2}+\Phi_2 x_{t-2}+Z_t
$$

$$
=\left(\Phi_1-I\right)\left(x_{t-1}-x_{t-2})+\left(\Phi_1+\Phi_2-I\right) x_{t-2}+Z_t \right.
$$
$$
=\left(\Phi_1-I\right) \Delta x_{t-1}+\left(\Phi_1+\Phi_2-I\right) x_{t-2}+Z_t .
$$
$$
\pi=\Phi_1+\Phi_2-1=\left(\begin{array}{cc}
\phi-1 & c-c \phi \\
0 & 1-1
\end{array}\right)=\left(\begin{array}{cc}
\phi-1 & c-c \phi \\
0 & 0
\end{array}\right)
$$
$$
\Gamma=\Phi_1-I=\left(\begin{array}{cc}
\Phi 1 & c \\
0 & 0
\end{array}\right)
$$



#Q2

#a
```{r,warning=FALSE}
library(AER)
library(urca)
data("PepperPrice")
log_p <- log(PepperPrice)
plot.ts(log_p)

```

#b
for black pepper
```{r}
ur.tr = ur.df(log_p[,1],type='trend')
summary(ur.tr)

```
Because -2.0712>-3.98, 2.3821<phi3, we fail to reject H0. So we conclude pi = 0 and beta2 = 0. Then, we continue the test.
```{r}
ur.dr2 = ur.df(log_p[,1],type='drift')
summary(ur.dr2)
```

-2.1721>tau2 and 2.4731<phi1, we fail to reject H0. So we conclude pi = 0 and beta1 = 0. Then, we continue the test.

```{r}
ur3 = ur.df(log_p[,1],type='none')
summary(ur3)

```
0.3486>tau1, we fail to reject H0. So we conclude pi = 0.It has a unit root and it is not stationary.

For white pepper
```{r}
ur.tr2 = ur.df(log_p[,2],type='trend')
summary(ur.tr2)

```
-1.782>tau3, 1.6134<phi3, we fail to reject H0. So we conclude pi = 0 and beta2 = 0. Then, we continue the test.
```{r}
ur.dr = ur.df(log_p[,2],type='drift')
summary(ur.dr)

```
-1.7855>tau2, 1.7474<phi1, we fail to reject H0. So we conclude pi = 0 and beta1 = 0. Then, we continue the test.
```{r}
ur.no2 = ur.df(log_p[,2],type='none')
summary(ur.no2)

```
0.4478>tau1, we fail to reject H0. So we conclude pi = 0. It has a unit root and it is not stationary

#c
```{r}
vecm.pepper = ca.jo(log_p)
summary(vecm.pepper)

```
We can say they are cointegrated. Rank 1 is suggested because we fail to reject r <= 1 and we reject r = 0 at 90% confident level.

#d
```{r}
vecm.r1 = cajorls(vecm.pepper,r = 1)
vecm.r1

```
black = 0.8904272 * white + error term


#Q3
```{r,warning=FALSE}
library(vars)
```

#(a)
```{r}

set.seed(1)
Ft = arima.sim(list(order = c(1,0,0),ar=0.7),n=50)
lambda = c(rep(0.25,10),rep(2,10))
err = rnorm(20*50,mean=0,sd=1)
err = matrix(err,nrow=20,ncol=50)
Yt = lambda%*%t(Ft) + err
par(mfrow = c(4,3))
par(mar=c(2,2,2,2))
for(l in (1:5)){
 plot.ts(Yt[l,])
}
for(l in (16:20)){
 plot.ts(Yt[l,])
}
ts.plot(Yt[c(1:5,16:20),], gpars = list(col=rainbow(10)))

```

#(b)

```{r}
source("DFM-VAR-library.R")
for(a in (1:20)){
 Yt[a,] = (Yt[a,] - mean(Yt[a,]))/sd(Yt[a,])
}
out = DFM.VAR(Yt, r=1)
plot.ts(t(out$Fh))
plot.ts(out$Lam,xlab="dim")

```
#(c)
```{r}
Ft_aver = colSums(Yt)/20
print(Ft_aver)
```
#(d)
```{r}

Ft_aver_standard = (Ft_aver-mean(Ft_aver))/sd(Ft_aver)
Ft_out = out$Fh[1,]
Ft_out_standard = (-1)*(Ft_out-mean(Ft_out))/sd(Ft_out)
Ft_standard = (Ft-mean(Ft))/sd(Ft)
plot(Ft_standard)
lines(Ft_aver_standard,col="red")
lines(Ft_out_standard,col="blue")
```
```{r}
print(mean(abs(Ft_standard - Ft_aver_standard)))
print(mean(abs(Ft_standard - Ft_out_standard)))

```

Both lines are in close proximity to the Ft line, but it appears that the standardized Ft_out line in blue fits marginally better. This is supported by the calculation of the mean absolute error, indicating that the Ft_out computed in part (b) is a better fit.



#(e)
```{r}
plot(out$Lam, type="l")
```

Equation (2) takes into account the various values of lambda. Simply averaging Yjt would not accurately measure F_t because it assumes all weights are equal. By using equation (2), the more important F_t in Y_t is weighted more heavily. Therefore, using the weights lambda as in equation (2) may work better than using equal weights as in equation (1).